### 2015-04-20 intro.tex
updates in abstract and intro

Data races lead to nondeterministically occurring bugs that can be hard to reproduce, isolate and debug, especially in the context of a complex operating system. \PDComment{maybe also mention that races in C lead to undefined behaviour and potential issues with compilers?} Linux kernel analysers, such as sparse~\cite{corbet2004sparse}, coccinelle~\cite{padioleau2008doc} and lockdep~\cite{corbet2006lock}, have successfully found deadlocks in kernel code, but are typically unable to detect data races. Techniques such as~\cite{ball2006thorough, clarke2004predicate, qadeer2004kiss, engler2000checking, henzinger2002temporal, cook2006termination, pratikakis2006locksmith, kuznetsov2010testing, renzelmann2012symdrive, lal2012corral} have been used to analyse the program behaviour of Linux and Windows device drivers, but primarily focus on reasoning about generic sequential program properties and finding protocol bugs. The few of these techniques that attempt to deal with data races are usually either \emph{unsound} (i.e. can miss real bugs) or \emph{imprecise} (i.e. can report false bugs). Furthermore, most previous techniques that attempt to reason about thread interleavings face significant scalability issues, because of the exponentially large state-space of realistic concurrent programs~\cite{musuvathi2008finding}.

In this paper, we present \whoop, a novel infrastructure for (i) automatically analysing the concurrent behaviour of Linux drivers written in the C language and (ii) accelerating external bug-finders that can easily plug in our toolchain. \whoop is empowered by \emph{static pairwise lockset analysis}, a novel technique for automatically verifying the absence of data races in Linux drivers (e.g. device drivers and filesystems). The key idea behind our approach is that a driver can be proven free from data races by (i) deriving a sound \emph{sequential} model that \emph{over-approximates} the originally concurrent program, (ii) instrumenting it for lockset analysis and race checking, and (iii) asserting that all accesses to the same shared resource are protected by at least one common lock. The immediate benefit is that our approach not only avoids reasoning about thread interleavings, and thus has the potential to scale well, but also allows the reuse of existing successful sequential verification techniques.

We then show that by exploiting the race-freedom guarantees that our static analysis provides, we can achieve a sound partial-order reduction that significantly accelerates \corral~\cite{lal2012solver}, a state-of-the-art bug-finder that is used by Microsoft to analyse device drivers and other concurrent programs. Using \whoop and \corral we analysed 10 drivers available in the Linux distribution.

>>>>

Data races lead to nondeterministically occurring bugs that can be hard to reproduce, isolate and debug, especially in the context of a complex operating system. According to the C language standard, data races also lead to undefined behaviour, which can be the source of various compilation and optimisation issues \PDComment{which is the best cite here?}. Linux kernel analysers, such as sparse~\cite{corbet2004sparse}, coccinelle~\cite{padioleau2008doc} and lockdep~\cite{corbet2006lock}, have successfully found deadlocks in kernel code, but are typically unable to detect data races. Techniques such as~\cite{ball2006thorough, clarke2004predicate, qadeer2004kiss, engler2000checking, henzinger2002temporal, cook2006termination, pratikakis2006locksmith, kuznetsov2010testing, renzelmann2012symdrive, lal2012corral} have been used to analyse the program behaviour of Linux and Windows device drivers, but primarily focus on reasoning about generic sequential program properties and finding protocol bugs. The few of these techniques that attempt to deal with data races are usually either \emph{unsound} (i.e. can miss real bugs) or \emph{imprecise} (i.e. can report false bugs). Furthermore, most previous techniques that attempt to reason about thread interleavings face significant scalability issues, because of the exponentially large state-space of realistic concurrent programs~\cite{musuvathi2008finding}.

In this paper, we present \whoop, a new approach that (i) statically analyses the concurrent behaviour of Linux device drivers written in the C language and (ii) uses the results from the static analysis to accelerate bug-finding. \whoop is empowered by \emph{static pairwise lockset analysis}, a novel technique for automatically verifying the absence of data races in Linux drivers. The key idea behind our approach is that a driver can be proven free from data races by (i) deriving a sound \emph{sequential} model that \emph{over-approximates} the originally concurrent program, (ii) instrumenting it for lockset analysis and race checking, and (iii) asserting that all accesses to the same shared resource are protected by at least one common lock. The immediate benefit is that our approach not only avoids reasoning about thread interleavings, and thus has the potential to scale well, but also allows the reuse of existing successful sequential verification techniques.

We then show that by exploiting the race-freedom guarantees that our static analysis provides, we can achieve a sound partial-order reduction that significantly accelerates \corral~\cite{lal2012solver}, a state-of-the-art bug-finder that is used by Microsoft to analyse device drivers, as well as other concurrent programs. Using \whoop and \corral we analysed \sizeOfBenchmarks drivers available in the 4.0 Linux kernel distribution.
